---
title: "Eli Boahen | Final Project: Cyborg Sensibilities"
author: "Eli Boahen"
description: "A survey of viewpoints on cybrgs and robots by Eli"
date: "07/02/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Final Project
  - Eli Boahen
  - RMarkdown Basics
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)
library(ggplot2)
library(ggrepel) 

knitr::opts_chunk$set(echo = TRUE)
```
This is my final project for DACSS 601. Thank you for all the support I received.

# Project Overview
This summer I release a survey exploring the current sentiments towards robots and cyborgs. Cyborgs and human like robots---or robot like humans---has been seen as a more Science Fiction idea and topic, however, I feel that with the increased eyes on A.I. and future-tech, these concepts are becoming more mainstream. This summer I also took an Internet Law and Policy course. I takeaway I had, was how US law tends to be reactive than proactive. This means laws tend to be made or drafted after an event or situation arises, not preemptively. I also noticed, that sentiments at the time of signing or creation play a huge role as to whether or not laws get passed, drafted, vetoed, or protested. The reactive nature of US law and the tech industry made me want to explore sentiments.

Sentiments can help us understand the populous opinion. This can tell us how individuals might vote, their concerns, their hopes, and their pain points. Through surveying, we can get a glimpse of the ways current opinions of robots and cyborgs, and forecast what policies might align with the populous. This is cyborg Anthropology.

## What is Cyborg Anthropology?
Cyborg Anthropology goes back over a decade, exploring Human Computer Interaction but from the lens of culture, beliefs, languages, and more. Think of mobile phones becoming phone vs smart phone, becoming smartphone and dumbphone. We have language like "a brick" being used to describe phones from the 90's and early 2000's, "a flip" phone versus "the flip" phone, and "indestructible" is more associated with early models of mobile phones that our modern devices. Cyborg Anthropology is the exploration of Human-ness and Robot-ness, and how we discuss and understand it.

## Data Overview
```{r Data Overview}
css_data <- read.csv("CSS_Responses_2023_admin.csv")
#head(css_data) <- I didn't get too much out of this
view(css_data)
summary(css_data)
```

The survey consists of 42 questions: 4 consent questions, 4 demographic information questions, and 34 sentiment related questions. For this project, I will focus on Section 2 ( questions 13 to 31 ) and Section 4 ( questions 32 to 39 ), also the interesting results of `"Are you a human?"`.
In this data

## Research Questions
>
- Currently, what does society believe to be the line between human and robot?
- Currently, what does society believe to be the line between human and cyborg?
- What is the overall sentiment towards robots?
- What is the overall sentiment towards cyborgs?
- When given cases that ask participants to apply their beleifs, do sentiments change?
> 
 
I entered this project with 3 of these questions, and gained the other 2 while writing the survey. I had a smooth 3 question goal but, in a larger research goal, I would love to explore what we believe is the line between robots, humans, and cyborgs. Many in the field of Cyborg Anthropology believe we are already cyborgs. Though that may work in a research and conceptual model, the daily life of our current society does not reflect that sentiment---neither does US policy. The fact is, the *feelings* we have about technology, robots, and cyborgs weigh more than the reality and definitions we have laied. Ironically, those feeling guide the definitions we put into law.

## Hypotheses
- The majority of participants will have a positive association with humans.
- The majority of participants will have a neutral association with cyborgs.
- The majority of participants will have a negative association with robots.
- 50% of participant sentiments will change after the survey. 


## Analysis


### Positive or Negitive Feelings
This was a learning process.

First, I attempted to analyze the text responses, however this proved to be a bigger challenge. I have added these attempts because there are still interesting observations to be drawn from said data, and I learned a lot along the way.

Second, I used columns 15 to 32 to get an idea of sentiments held by participants

```{r Friendship with Other-Beings Data Frame}
#colnames(css_data)

# Friendship with a Cyborg
cyborg_friendship <- table(css_data[,15])
cyborg_friendship

cyborg_friendship_df <- as.data.frame(cyborg_friendship)
names(cyborg_friendship_df)[1] <- "Response"
cyborg_friendship_df

# friendship with a Robot
robot_friendship <- table(css_data[,16])
robot_friendship

robot_friendship_df <- as.data.frame(robot_friendship)
names(robot_friendship_df)[1] <- "Response"
robot_friendship_df
```
```{r Friendship Graphs}
# Cyborg Friendship Graph
cyborg_friendship_graph <- ggplot(data = cyborg_friendship_df,
                            aes(x = Response,
                                y = Freq,
                                fill = "steelblue")
                            ) +
  geom_bar(stat="identity"
           )+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
cyborg_friendship_graph

# Robot Friendship Graph
robot_friendship_graph <- ggplot(data = robot_friendship_df,
                            aes(x = Response,
                                y = Freq,
                                fill = "darkturquoise")
                            ) +
  geom_bar(stat="identity"
           )+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
robot_friendship_graph
# Do I know why it's pink? No. Does it look pretty? Yes.
```


```{r Owning Other-Beings Data Frame}
#colnames(css_data)

# owning a Cyborg
cyborg_owning <- table(css_data[,17])
cyborg_owning

cyborg_owning_df <- as.data.frame(cyborg_owning)
names(cyborg_owning_df)[1] <- "Response"
cyborg_owning_df

# owning a Robot
robot_owning <- table(css_data[,18])
robot_owning

robot_owning_df <- as.data.frame(robot_owning)
names(robot_owning_df)[1] <- "Response"
robot_owning_df
```
```{r Owning Graphs}
# Cyborg Owning Graph
cyborg_owning_graph <- ggplot(data = cyborg_owning_df,
                            aes(x = Response,
                                y = Freq
                                )
                            ) +
  geom_bar(stat="identity",
           fill = "steelblue"
           )+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
cyborg_owning_graph

# Robot Owning Graph
robot_owning_graph <- ggplot(data = robot_owning_df,
                            aes(x = Response,
                                y = Freq
                                )
                            ) +
  geom_bar(stat="identity",
           fill = "darkturquoise"
           )+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
robot_owning_graph
# Figured it out: define fill in geom_bar, not in ggplot
```

The data goes on, but overall, the charts mirror each other. From friendship to rights, participants leaned in favor of cyborgs and away from robots. The few times sentiments for cyborgs and for robots aligned was column `27` and `28` as well as `31` with`32`. The majority answered `No`. This is strange because for columns `17` and `18`, 76.47% of respondents answered "Yes" that humans should be able to own a robot. Yet, column `26` and `27`, 70% of participants said they personally would not own a robot. This is ironic given how much technology the average human has.


There seems to be a much more positive view of cyborgs that I expected.


#### The Exploration of Data and Text
```{r Creating Frequency of Feelings Table}
# Sources: https://www.youtube.com/watch?v=n7N1dFTj5Lc
#         https://stackoverflow.com/questions/7531868/how-to-rename-a-single-column-in-a-data-frame


# First seeing what this data looks like
freq_feels <- sort(table(unlist(strsplit(css_data$A.human.is......, " "))), 
               decreasing = TRUE)
freq_feels

# Giving the data a scrub down
# Attempted . . . switching to manually

# Next, turning it into a data frame. This was fun because I half guessed it, then searched online for support.
# This chunky Chunk, takes the `A.human.is` column, replaces all non-letters with a space splits them by the space, sorts, then turns them into a table. Very chnky indeed
freq_feels <- sort(
                  table(
                        unlist(
                              strsplit(
                                      str_replace_all(
                                          tolower(css_data$A.human.is......), "[^[:alnum:]]", " "), " ")
                              )
                        )
                  )
# Turns all of that into a data frame
freq_feels_human_df <- as.data.frame(freq_feels)
freq_feels_human_df

# Let's Rename the columns
names(freq_feels_human_df)[1] <- "Words"
names(freq_feels_human_df)[2] <- "Frequency"
freq_feels_human_df

#Vuala!
view(freq_feels_human_df)

```

Many lines of code were learned through in the production of this analysis, ultimately, this may not be the ideal presentation, but I think we can learn from it.

```{r An Attempt Was Made}
# Attempting Cleaning Again

# Tried to go this route, but decided to go with the raw data instead of preemptively deciding which is valuable
#freq_feels_human_subset_css <- freq_feels_human_subset[c(5,7,9,11,),]

# Did not work
#freq_feels_human_subset_css <- freq_feels_human_subset[1,] + freq_feels_human_subset[4,]
#freq_feels_human_subset_css
```

Repeating that for cyborgs and robots
```{r Cyborg Frequency of Feelings}

# The chunky Chunk
cyborg_freq_feels <- sort(
                  table(
                        unlist(
                              strsplit(
                                      str_replace_all(
                                          tolower(css_data$A.robot.is......), "[^[:alnum:]]", " "), " ")
                              )
                        )
                  )
# Turns all of that into a data frame
freq_feels_cyborg_df <- as.data.frame(cyborg_freq_feels)
freq_feels_cyborg_df

# Let's Rename the columns
names(freq_feels_cyborg_df)[1] <- "Words"
names(freq_feels_cyborg_df)[2] <- "Frequency"
freq_feels_cyborg_df

#Vuala!
view(freq_feels_cyborg_df)

```

```{r Robots Frequency of Feelings}

# The chunky Chunk
robot_freq_feels <- sort(
                  table(
                        unlist(
                              strsplit(
                                      str_replace_all(
                                          tolower(css_data$A.robot.is......), "[^[:alnum:]]", " "), " ")
                              )
                        )
                  )
# Turns all of that into a data frame
freq_feels_robot_df <- as.data.frame(robot_freq_feels)
freq_feels_robot_df

# Let's Rename the columns
names(freq_feels_robot_df)[1] <- "Words"
names(freq_feels_robot_df)[2] <- "Frequency"
freq_feels_robot_df

#Vuala!
view(freq_feels_robot_df)

```

**Graphing Results**
Here was my first trial at graphing.
```{r First Plot}
ggplot(data = freq_feels_human_df,
       mapping = aes(x = Words,
                     y = Frequency))+
  geom_point(size = 1)
```

Here are my final charts
```{r Human Word Map Graph}
freq_feels_human_subset <- filter(freq_feels_human_df, Frequency >= 2 )
freq_feels_human_subset
human_feels_graph <- ggplot(data = freq_feels_human_subset,
                            aes(x = Words,
                                y = Frequency,
                                fill = Frequency)
                            ) +
  geom_bar(stat="identity",
           aes(color = Frequency
               )
           )+
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE)
           )+
  scale_fill_gradient(low = "midnightblue", high = "darkorchid1")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
human_feels_graph
```

```{r Cyborg Word Map Graph}
freq_feels_cyborg_subset <- filter(freq_feels_cyborg_df, Frequency >= 2 )
freq_feels_cyborg_subset
cyborg_feels_graph <- ggplot(data = freq_feels_cyborg_subset,
                            aes(x = Words,
                                y = Frequency,
                                fill = Frequency)
                            ) +
  geom_bar(stat="identity",
           aes(color = Frequency
               )
           )+
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE)

           )+
  scale_colour_gradient(low = "midnightblue", high = "steelblue")

cyborg_feels_graph
```

```{r Robot Sentiments}
freq_feels_robot_subset <- filter(freq_feels_robot_df, Frequency >= 2 )
freq_feels_robot_subset
robot_feels_graph <- ggplot(data = freq_feels_robot_subset,
                            aes(x = Words,
                                y = Frequency,
                                fill = Frequency)
                            ) +
  geom_bar(stat="identity",
           aes(color = Frequency
               )
           )+
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE)

           )+
  scale_colour_gradient(low = "midnightblue", high = "darkturquoise")

robot_feels_graph
```


With these graphs, we can peer into some of the word associations towards `humans`, `cyborgs`, and `robots`

### Sentiment Change
This question asks the participants if they're feelings have changed. It comes at the closing of the survey, after participants respond to 3 human-cyborg-robot related questions.

In which direction did said sentiments change? That is unknown. However, this data can tell us how sentiments towards robots or cyborgs can change *after* one faces the application of their morals. I wanted to explore this question, because it is itself an experiment, one can hold one sentiment, but apply another; it's a reflection of our current reactive model in policy and industry.

```{r Sentiment Changes Data Frame}
# Taking column 41, "Did your sentiments change", turning it into a table, which gives us the frequency
sentiment_changes <- table(css_data[,41])
sentiment_changes

# Then we turn it into a data frame
sentiment_changes_df <- as.data.frame(sentiment_changes)
names(sentiment_changes_df)[1] <- "Sentiment_Shift"
sentiment_changes_df

# Check your results
summary(sentiment_changes_df)
```
```{r Sentiment Percent Change}
summary(sentiment_changes_df)
# Doing this manually
#(sum(sentiment_changes_df$Freq))/sum(sentiment_changes_df$Freq)

a_change_in_sentiment <- ((6+6+3)/17)*100
min_some_change_in_sentiment <- ((6+6)/17)*100
full_change_in_sentiment <- ((3)/17)*100
no_change_in_sentiment <- ((2)/17)*100
a_change_in_sentiment
min_some_change_in_sentiment
full_change_in_sentiment
no_change_in_sentiment

```

We can see, the majority of participants had some change in their held sentiment at the start of the survey. 88.23% had some amount of change while 11.76% had a full change in sentiments

The majority participants had a minimal or slight shift in sentiment. This shows and proves that facing cases, could lead to a sentiment shift in the populous.

```{r Sentiment Graph}
sentiment_graph <- ggplot(data = sentiment_changes_df,
                            aes(x = Sentiment_Shift,
                                y = Freq,
                                fill = Freq)
                            ) +
  geom_bar(stat="identity"
           )+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
  
sentiment_graph
```

### Am I A Human?
```{r}
# This was the first one I worked through, so there are significantly more comments and tests
# colnames(css_data)
# css_data$Are.you.a.human. <- Goog for checking that I get the right result with the code below
t_human_ness_subset_css <- css_data[2:17,8] # "t_" for testing
t_human_ness_subset_css

human_ness_subset_css <- css_data[,8]# even though this and t_human_ness_subset_css are the same, I experimented with the test variable until I got a good answer / format
human_ness_subset_css
```



## Results

### Hypotheses Revisted
X for False
✓ for True
- [X] The majority of participants will have a positive association with humans.
- [X] The majority of participants will have a neutral association with cyborgs.
- [✓] The majority of participants will have a negative association with robots.
- [X] 50% of participant sentiments will change after the survey. 

#### The majority of participants will have a positive association with humans.
The response towards humans was generally neutral. based on the words used to describe humans, the response was quite neutral. It assumes many participants would answer more cynically, however the responses were quite scientific as well as conceptual---the concept of a human. There was one negative resoponse i the word association section.

I feel that a survey with direct questions about humans and humanness would be better suited for a hypothesis like this, however I think that gets more into morality and may distract from the overall survey.

#### The majority of participants will have a neutral association with cyborgs.
Actually, the responses leaned towards positive. I believe this is influende by media as film and media were referenced in participant responses.

#### The majority of participants will have a negative association with robots.
True. Participants were not in favor of friendhip, nor rights for robots. In addition, the word association section for robots contained words such as "scary", "cold", and frequently describing it's materiality rather than a capability---humans had "emotions", "brain", and "biology" while robots had "machine", "metal", and "computer".

#### 50% of participant sentiments will change after the survey.
The data shows that 88.23% had a change in sentiment. My hypothesis was wrong; signifigaly more participants had a change that I predicted.

# Conclusion
Internet law started as a proactive step. Laws like Section 230 made the Internet what it is today, and it was done by collaborative efforts. Few could have foreseen the full realized potential of the internet, yet Chris Cox and Ron Wyden foresaw the Internet's impact on Free Speech. When proactive exploration and collaboration happens, we get the ability to react with guardrails versus attempting to wrangle as a situation spirals further.

With this survey, I hope to take the first step of many towards understanding human and non-human sentiment and culture to make societal and policy change.

Best wishes, 
Eli
 