---
title: "Eli Boahen | Tutorial 3"
author: "Eli Boahen"
description: "A run through of tutorial 3"
date: "07/06/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Tutorial3
  - Eli Boahen
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```
In this tutorial we talk about all things data and `R`. How to read in data, ___, ___, and _____

First, let's kick off with the glossary from tutorial 2. I've edited the origial descriptions & added some of my own.

## Helpful Commands
Tutorial 1: R Basics and Assigning Variables

- Math: `+`,`-`,`*`,`/`
- Booleans & Logical Operators: `==`,`>`,`<`,`>=`,`>=`
- Assigning variables: `variable_name <- "value"`
- Create vector ( the `R` version of lists ): `c()`

Tutorial 2: Reading and Describing Data

- Preview data: `head()`
- Get dimensions: `dim()`
- Get column names and numbers: `colnames()`
- Select specific columns: `select()`
- `starts_with()`
- `ends_with()`
- `contains()`
- Get values of column: `table()`
- Get proportions of values of column: `prop.table()`
- Tidy data: `gather()` and spread()`

Tutorial 3
- `pivot_longer()`
- `pivot_wider()`

## What is Data?
Dataframes, or `df`, are 2-dimensional data structures--so tables



```{r}
x <- c(2,3,4,5)
mean(x)
```

Let's introduce the 3 data-horsemen of our Tutorial 3:
`pseo`, `electricity`, `anes`
---------------------------------------------
## Rendering your post

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. 

:::{.callout-warning}
Be sure that you have moved your `*.qmd` file into the `posts` folder **BEFORE** you render it, so that all files are stored in the correct location.
:::

:::{.callout-important}
Only render a single file - don't try to render the whole website!
:::

:::{.callout-note}
## Pilot Student Blogs

We are piloting a workflow including individual student websites with direted and limited pull requests back to course blogs. Please let us know if you would like to participate.
:::

## Reading in data files

The easiest data source to use - at least initially - is to choose something easily accessible, either from our `_data` folder provided, or from an online source that is publicly available.

:::{.callout-tip}
## Using Other Data
If you would like to use a source that you have access to and it is small enough and you don't mind making it public, you can copy it into the `_data` file and include in your *commit* and *pull request*.
:::

:::{.callout-tip}
## Using Private Data

If you would like to use a proprietary source of data, that should be possible using the same process outlined above. There may initially be a few issues. We hope to have this feature working smoothly soon!
:::