---
title: "Eli Boahen | Tutorial 5"
author: "Eli Boahen"
description: "Tutorial 5 Completion"
date: "07/13/2023"
format:
  html:
    toc: true
    code-fold: true
    code-copy: true
    code-tools: true
categories:
  - Tutorial5
  - EliBoahen
---

```{r}
#| label: setup
#| warning: false

library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

## What did you learn?
We learned what is piping and how to use `%>%`, `group_by()`, `function(){}`, and `for(){}`

## What did you remember?
Order matters a lot for piping

## Anything you didn't understand?
I dont' reeaaaly understand why piping isn't on every line of R
```{r}
# Later on I thought of it like Mario:
# Going into a pipe, doing tasks & exploring a world
```

## Something you would change?
The *Writing Functions* section is a bit long and intimidating with the uncollapsible R code

Not is a mean way at all! I saw a typo in Piping basics, second to last paragraph, "oten". I make the same mistake myself and often miss or don't flag it because "f" and "t" look similar to me.


Best wishes,
Eli